---
title: "p8105_hw1_by2287"
author: "Binyam Yilma"
date: "2020-09-15"
output: github_document
---

```{r load libraries, include=FALSE}
library(tidyverse)
```

## Problem 1

Create a data frame named 'df' comprised of:
```{r}
df <- tibble(
  vec_numeric = rnorm(10),
  vec_logical = vec_numeric < 0,
  vec_character = c("a", "b", "c", "d", "e", "f", "g", "h", "i", "j"),
  vec_factor = factor(c("x", "y", "z", "x", "y", "z", "x", "y", "z", "x")),
)
```

#####Taking the mean of each variable in the data frame "df"

#####Taking the mean of the numeric & logical variable returns a value; but only an error message for the character & factor variables
```{r}
#taking the mean of the numeric variable
df %>% 
  pull("vec_numeric") %>% 
  mean()
#taking the mean of the logical variable
df %>% 
  pull("vec_logical") %>% 
  mean()
#taking the mean of the character variable
df %>% 
  pull("vec_character") %>% 
  mean()
#taking the mean of the factor variable
df %>% 
  pull("vec_factor") %>% 
  mean()

```

The as.numeric function coerces the non-numeric variables in "df" into numeric ones. The logical variable turned to 0 & 1; whereas the factor variable variable truned to 1, 2, & 3; the character variable turned into NAs. This explains what we saw when we took the mean for the non-numeric variables. It appears that R reads the logical values as 0 or 1; and the factor variable variables as 1, 2, & 3 - and thus takes the average of these numbers hen we call the mean() function. 
```{r}
as.numeric(df$vec_logical)
as.numeric(df$vec_character)
as.numeric(df$vec_factor) 
```


Converting the logical vector to different variable types and multiplying by the random sample (vec_numeric)
```{r}
#converting to numeric and multiplying by the random sample
df$vec_logical <- as.numeric(df$vec_logical)
pull(df, "vec_logical") * pull(df, "vec_numeric")

#converting to factor and multiplying by the random sample
df$vec_logical <- as.factor(df$vec_logical)
pull(df, "vec_logical") * pull(df, "vec_numeric")

#1st converting to factor
#then converting to numeric then
#multiplying by the random sample
df$vec_logical <- as.factor(df$vec_logical)
df$vec_logical <- as.numeric(df$vec_logical)
pull(df, "vec_logical") * pull(df, "vec_numeric")
```


## Problem 2
```{r}
library("palmerpenguins")
data("penguins", package = "palmerpenguins")
```


**Description of the data set Penguins**

The data set _penguins_ has `r nrow(penguins)` rows and `r ncol(penguins)` variables, which capture different types of information on penguins, including: `r names(penguins)`. 

The data set captures information on 3 different species of penguins: `r levels(penguins$species)`; across 3 different islands: `r levels(penguins$island)`.

The median bill length of these penguins, not counting missing values, is `r pull(penguins, bill_length_mm) %>% median(na.rm = TRUE)` millimeter; whereas the median bill depth is `r pull(penguins, bill_depth_mm) %>% median(na.rm = TRUE)` millimeter.

The average flipper length of these penguins, not counting missing values, is `r pull(penguins, flipper_length_mm) %>% mean(na.rm = TRUE) %>% round(2)` millimeter. 


